Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-02-05 article.tex
Proof of Proposition 1

A \emph{bivincular pattern} (abbreviated BVP) $\sigma$
of length $k$ is a permutation in $S_k$ written in
two-line notation 
(that is the top row is $12\,\ldots\,k$ and the bottom row
is a permutation $\sigma_1\sigma_2\,\ldots\,\sigma_k$).
We have the following conditions on the top and bottom rows
of $\sigma$:
\begin{itemize}
	\item
	If the bottom line of $\sigma$ contains
	$\underline{\sigma_i\sigma_{i+1}\,\ldots\,\sigma_j}$
	then the letters corresponding to 
	$\sigma_i\sigma_{i+1}\,\ldots\,\sigma_j$ in an occurrence of
	$\sigma$ in a permutaton must be adjacent, whereas there is
	no adjacency condition for
	non-underlined consecutive letters.
	Moreover if the bottom row of $\sigma$ begins with
	$_\llcorner{\sigma_1}$ then any occurrence of $\sigma$ 
	in a permutation $\pi$ must begin with the leftmost 
	letter of $\pi$,
	and 
	if the bottom row of $\sigma$ begins with
	${\sigma_k}_\lrcorner$ then any occurrence of $\sigma$ 
	in a permutation $\pi$ must end with the rightmost 
	letter of $\pi$.
	\item
	If the top line of $\sigma$ contains
	$\overline{i\,i+1\,\ldots\,j}$ then the letters corresponding to
	$\sigma_i, \sigma_{i+1}, \ldots, \sigma_j$ in an
	occurrence of $\sigma$ in a permutation must be adjacent in values,
	whereas there is no value adjacency restriction for non-overlined 
	letters.
	Moreover, if the top row of $\sigma$ begins with
	$^\ulcorner{1}$ then
	any occurrence of $\sigma$ is a permutation $\pi$ must begin with
	the smallest letter of $\pi$, and
	if top row of $\sigma$ ends with $k^\urcorner$ then
	any occurrence of $\sigma$ is a permutation $\pi$ must end with
	the largest letter of $\pi$.
\end{itemize}
A bivincular pattern avoids a set of permutations $P$ if its bottom
row $\sigma_1\sigma_2\,\ldots\,\sigma$ avoids every permutation in $P$.

@@

\begin{proof}
To solve the permutation pattern for a permutation in $\AV_n(213,231)$, 
we find a find a match of the stripe of the pattern in the stripe of the text.
This can be done linearly by a greedy algorithm, 
by matching a step whenever it is possible.
And thank to the definition of the definition of a stripe in \textbf{Corollary \ref{corollary:minmaxelement}},
we do not have to compute the stripe in advance.
Therefore we have an linear (online) algorithm.
\end{proof}

>>>>

A \emph{bivincular pattern} (abbreviated BVP) $\sigma$
of length $k$ is a permutation in $S_k$ written in
two-line notation 
(that is the top row is $12\,\ldots\,k$ and the bottom row
is a permutation $\sigma_1\sigma_2\,\ldots\,\sigma_k$).
We have the following conditions on the top and bottom rows
of $\sigma$:
\begin{itemize}
	\item
	If the bottom line of $\sigma$ contains
	$\underline{\sigma_i\sigma_{i+1}\,\ldots\,\sigma_j}$
	then the letters corresponding to 
	$\sigma_i\sigma_{i+1}\,\ldots\,\sigma_j$ in an occurrence of
	$\sigma$ in a permutaton must be adjacent, whereas there is
	no adjacency condition for
	non-underlined consecutive letters.
	Moreover if the bottom row of $\sigma$ begins with
	$_\llcorner{\sigma_1}$ then any occurrence of $\sigma$ 
	in a permutation $\pi$ must begin with the leftmost 
	letter of $\pi$,
	and 
	if the bottom row of $\sigma$ begins with
	${\sigma_k}_\lrcorner$ then any occurrence of $\sigma$ 
	in a permutation $\pi$ must end with the rightmost 
	letter of $\pi$.
	\item
	If the top line of $\sigma$ contains
	$\overline{i\,i+1\,\ldots\,j}$ then the letters corresponding to
	$\sigma_i, \sigma_{i+1}, \ldots, \sigma_j$ in an
	occurrence of $\sigma$ in a permutation must be adjacent in values,
	whereas there is no value adjacency restriction for non-overlined 
	letters.
	Moreover, if the top row of $\sigma$ begins with
	$^\ulcorner{1}$ then
	any occurrence of $\sigma$ is a permutation $\pi$ must begin with
	the smallest letter of $\pi$, and
	if top row of $\sigma$ ends with $k^\urcorner$ then
	any occurrence of $\sigma$ is a permutation $\pi$ must end with
	the largest letter of $\pi$.
\end{itemize}
% A bivincular pattern avoids a set of permutations $P$ if its bottom
% row $\sigma_1\sigma_2\,\ldots\,\sigma$ avoids every permutation in $P$.

@@

\begin{proof}
According to Lemma~\ref{lemma:MatchStripeToPermutation} the problem reduces 
to deciding whether $s_\sigma$ occurs as a subsequence in $s_\pi$. 
A straightforward greedy approach solves this issue in linear-time. 
Furthermore, in the light of Corollary~\ref{corollary:minmaxelement}, 
the effective construction of the two stripes is not required, 
so it is an online algorithm.
\qed
% To solve the permutation pattern for a permutation in $\AV_n(213,231)$, 
% we find a find a match of the stripe of the pattern in the stripe of the text.
% This can be done linearly by a greedy algorithm, 
% by matching a step whenever it is possible.
% And thank to the definition of the definition of a stripe in 
% Corollary~\ref{corollary:minmaxelement},
% we do not have to compute the stripe in advance.
% Therefore we have an linear (online) algorithm.
\end{proof}
